---
title: "Chapter 4-5: Data Wrangling key"
author: "Math 4180"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Transformatino with `dplyr`.
This .Rmd file steps examples of how to use the five key functions from `dplyr`: `filter()`, `arrange()`, `select()`, `mutate()` and `summarize()`. For the examples, we will use the `flights` data from the `nycflights13` package, and `dplyr` functions that are a part of the `tidyverse` package. To get started, we need to load the libraries and let's also look at our data so we can readily see the variable names and types.

```{r load-libraries, message=F}
library(nycflights13)
library(tidyverse)
```

Let's create a subsets of your data by selecting rows and arranging rows. First, let's take a look at the data so that we know what information we are given:
```{r data_check}
??nycflights13::flights
flights %>% glimpse()
```

# Using `filter()`

The filter command is used to select rows based on values.

To select values, we will make use of the logical operators of `==` (equal), `&` (and), `|` (or), and `!` (not). We can use this in combination with `<`, `>`, `<=`, and `>=`.

Let's work through using `filter()` to find all flights that

1. Had an arrival delay of two or more hours (Ans. 10,200 rows)
```{r filter_1}
flights %>% 
  filter(arr_delay >= 120)
```

2. Flew to Houston (IAH or HOU) (Ans. 9,313 rows)
```{r filter_2}
flights %>%
  filter(dest == "IAH" | dest == "HOU")

flights %>%
  filter(dest %in% c("IAH", "HOU"))
```

3. Were operated by United, American, or Delta (Ans. 139,504 rows)
```{r filter_3}
flights %>% filter(carrier == "UA" | carrier == "AA" | carrier == "DL")

flights %>% filter(carrier %in% c("UA","AA","DL"))
```

4. Departed in summer (July, August, and September) (Ans. 86,326 rows)
```{r filter_4}
flights %>% filter(month == 7 | month == 8 | month == 9)
flights %>% filter(month %in% c(7, 8, 9))
flights %>% filter(month %in% 7:9)
```

5. Arrived more than two hours late, but did not leave late
```{r filter_5}
flights %>% filter(arr_delay > 120) %>% filter(dep_delay <= 0) 
flights %>% filter(arr_delay > 120 & dep_delay <= 0)
```

6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r filter_6}
flights %>% filter(dep_delay > 60) %>% filter(arr_delay <= 30)
flights %>% filter(dep_delay > 60 & arr_delay <=30)
```

7. Departed between midnight and 6 am (inclusive)
```{r filter_7}
flights %>% filter(dep_time <= 600 )
```

8. Use the `summary` command to find out where there are missing values in the `flights` dataset.

```{r filter_8, eval=F}
flights %>% summary()
```

# Using `arrange()`

The arrange command works similarly to `filter()` but instead of selecting rows, it reorders the rows in ascending order.

 1. Arrange the flights by departure time.
```{r arrange_1}
flights %>% arrange(dep_time)
```
 
 2. Are there missing values for departure time on any of the flights? Use `filter` and the `is.na` function.
```{r arrange_2}
flights %>% filter(is.na(dep_time))
```
 
 3. What if we did this using the `arrange` function.. where does `arrange()` place the rows with missing values?
 
```{r arrange_3}
flights %>% arrange(-is.na(dep_time))
```
 
 4. Use the `summary()` function to determine other variables that have missing values. 

```{r arrange_4}
flights %>% summary()
```

We can use the `desc()` command can be used within `arrange()` to reorder the elements in descending order. As an example, arrange the flights from December 1 to December 31. 

```{r arrange_desc}
flights %>%
  filter(month == 12) %>%
  arrange(desc(month, day))
```

# Using `select()`
While `filter()` and `arrange()` work on the rows of the data,  `select()` is used to select columns. Here are some examples:

```{r select_examples, eval=F}
flights %>%
  select(year, month, day)

flights %>%
  select(1, 2, 3)

flights %>%
  select(year:carrier)

flights %>%
  select(carrier, tailnum, everything())

flights %>%
  select(starts_with("a"))

flights %>%
  select(ends_with("e"))

flights %>%
  select(contains("th")) 

flights %>%
  select(matches("(.)\\1"))  #finds variables with repeated characters use ?matches

# other selects include all_of() or any_of()
```

To add new variables to a data set we use `mutate()`. Let's see if we can address these questions:


```{r mutate_example, eval=F}
flights %>%
  mutate(flight_time=distance/air_time) %>%
  select(flight_time, everything()) %>%
  arrange(flight_time)
```









