---
title: "Chapter 4-5: Data Wrangling"
author: "Math 4180"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Data Transformatino with `dplyr`.
This .Rmd file steps examples of how to use the five key functions from `dplyr`: `filter()`, `arrange()`, `select()`, `mutate()` and `summarize()`. For the examples, we will use the `flights` data from the `nycflights13` package, and `dplyr` functions that are a part of the `tidyverse` package. To get started, we need to load the libraries and let's also look at our data so we can readily see th variable names and types.

```{r cars, eval=F}
library(nycflights13)
library(tidyverse)
data(flights)  #adds the data set to the Global Environment
View(flights)  #opens a spreadsheet to view the data
```

##Creating a subset of your data by selecting rows and arranging rows. 

# Using `filter()` and `arrange()`
The filter command is used to select rows based on values.

To select values, we will make use of the logical operators of `==` (equal), `&` (and), `|` (or), and `!` (not). We can use this in combination with `<`, `>`, `<=`, and `>=`.

Let's work through using `filter()` to find all flights that

1. Had an arrival delay of two or more hours (Ans. 10,200 rows)
2. Flew to Houston (IAH or HOU) (Ans. 9,313 rows)
3. Were operated by United, American, or Delta (Ans. 139,504 rows)
4. Departed in summer (July, August, and September) (Ans. 86,326 rows)
5. Arrived more than two hours late, but didn’t leave late
6. Were delayed by at least an hour, but made up over 30 minutes in flight
7. Departed between midnight and 6 am (inclusive)

8. Use the `summary` command to find out where there are missing values in the `flights` dataset.

```{r filter, eval=F}
#1
flights %>%
  filter(arr_delay >=  )

```

The arrange command works similarly to `filter()` but instead of slecting rows, it reorders the rows in ascending order.
 1. Arrange the flights by departure time.
 2. Are there missing values for departure time on any of the flights? use the `is.na` function.
 3. Where does `arrange()` place the rows with missing values?
 4. Use the `summary()` function to determine other variables that have missing values. 

We can use the `desc()` command can be used within `arrange()` to reorder the elements in descending order. As an example, arrange the flights from December to January. 

```{r arrange, eval=F}




```


While `filter()` and `arrange()` work on the rows of the data,  `select()` is used to select columns. Here are some examples:

```{r, eval=F}
flights %>%
  select(year, month, day)

flights %>%
  select(1, 2, 3)

flights %>%
  select(year:carrier)

flights %>%
  select(carrier, tailnum, everything())

flights %>%
  select(starts_with("a"))

flights %>%
  select(ends_with("e"))

flights %>%
  select(contains("th")) 

flights %>%
  select(matches("(.)\\1"))  #finds variables with repeated characters use ?matches

# other selects include all_of() or any_of()
```

To add new variables to a data set we use `mutate()`. Let's see if we can address these questions:


```{r, eval=F}
flights %>%
  mutate(flight_time=distance/air_time) %>%
  select(flight_time, everything()) %>%
  arrange(flight_time)

```









