---
title: 'Data Vizualization using ggplot'
subtitle: "Part 2"
institute: "University of Minnesota Duluth"
date: "Spring 2021"
author: "MATH 4180"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

class:inverse, middle, center
#single categorical variable
---
class: middle

.pull-left[
If you could speak to R in English, how would you tell R to make this plot for you?

OR

If you had the data and had to draw this bar plot by hand, what would you do?
]

.pull-right[
```{r, echo=F}
library(ggplot2)
ggplot(iris) +
  aes(x = Species) +
  geom_bar()
```
]
---
class: middle

.pull-left[
Possible ideas

- Consider the data frame
- Count number of flowers in each **species**
- Put **species** on x-axis
- Put **count** on y-axis
- Draw the bars
]

.pull-right[
```{r, echo=F}
ggplot(iris) +
  aes(x = Species) +
  geom_bar()
```
]
---
class: middle

.pull-left[
These ideas are all correct but some are not necessary in R

- Consider the data frame
- ~~Count number of flowers in each **species**~~
- Put **species** on x-axis
- ~~Put **count** on y-axis~~
- Draw the bars

R will do some of these steps by default. Making a bar plot with another tool might look slightly different.
]

.pull-right[
```{r, echo=F}
ggplot(iris) +
  aes(x = Species) +
  geom_bar()
```
]
---
class: middle

# 3 steps of making a basic ggplot

1. Pick dataset

2. Map features onto aesthetics

3. Add the geometric layer
---
# 1. pick dataset

.pull-left[
```{r, eval=F}
* ggplot(iris)
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)
```
]
---
# 2. Map features onto aesthetics

.pull-left[
```{r, eval=F}
ggplot(iris) +
*  aes(x = Species)
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)+
  aes(x=Species)
```
]
---
# 3. Add the geometric layer

.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x = Species) +
*  geom_bar()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)+
  aes(x=Species)+
  geom_bar()
```
]
---
# Overview: Single Categorical Variable

1. Create a ggplot using the **iris** data frame.
2. Map the **Species** to the x-axis.
3. Add a layer of a bar plot.

.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x = Species) +
  geom_bar()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)+
  aes(x=Species)+
  geom_bar()
```
]
---
class:inverse, middle, center
#Single Numeric Variable
---
# Single Numerical Variable

.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)+
  aes(x=Sepal.Length)+
  geom_histogram()
```
]
---
# 1. Pick dataset

.pull-left[
```{r, eval=F}
* ggplot(iris)
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)
```
]
---
# 2. Map features onto aesthetics

.pull-left[
```{r, eval=F}
ggplot(iris) +
*  aes(x = Sepal.Length)
```
]
.pull-right[
```{r, echo=F}
ggplot(iris)+
  aes(x=Sepal.Length)
```
]
---
# 3. Add the geometric layer

.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
*  geom_histogram()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram()
```
]
---
#What is this warning?
```{r, echo=F, out.height=400}
ggplot(iris)+
  aes(x = Sepal.Length) +
  geom_histogram()
```

---
#binwidth=5

.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x=Sepal.Length) +
*  geom_histogram(binwidth = 5)
```
]
.pull-right[
```{r, echo=F, out.height=400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 5)
```
]
---
#binwidth=2
.pull-left[
```{r, eval=F}
ggplot(iris)+
  aes(x = Sepal.Length) +
*  geom_histogram(binwidth = 2)
```
]
.pull-right[
```{r, echo=F, out.height=400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 2)
```
]

---
#binwidth=0.5

.pull-left[
```{r, eval=F}
ggplot(iris)+
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5)
```
]
.pull-right[
```{r, echo=F, out.height=400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5)
```
]
---
#binwidth=1
.pull-left[
```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
*  geom_histogram(binwidth = 1)
```
]
.pull-right[
```{r, echo=F, out.height=400}
ggplot(iris)+
  aes(x=Sepal.Length)+
  geom_histogram(binwidth=1)
```
]

---
class: middle
[There is no "best" number of bins](https://en.wikipedia.org/wiki/Histogram#Number_of_bins_and_width)
---
class: middle 

.pull-left[
**Experiment!**


```{r, eval=F}
sample(iris$Sepal.Length, 10)
count(iris$Sepal.Length)
min(iris$Sepal.Length)
max(iris$Sepal.Length)
table(iris$Sepal.Length)
plot(iris$Sepal.Length)
```

]

--
.pull-right[
See which method you prefer to use. Do you prefer tables? Subsets? Outliers? Plots? 

Visualizing these relationships is usually the better option, but not always! Some graphics may be hard to interpret. Other times we may want to know exact values -- this is where a table can really shine.

]
---
class: middle

Pick your favorite [colors](/slides/Rcolor.pdf)
---
class:middle


```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
*                  fill = "coral3")
```

.pull-right[
```{r, echo=F, out.height=400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
                 fill = "coral3")
```
]
---
class: middle


```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
*                  color= "goldenrod4")
```

.pull-right[
```{r, echo=F, out.height = 400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
                 color = "goldenrod4")
```
]
---
class: middle


```{r, eval=F}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
*                  fill = "coral3",
*                  color = "goldenrod4")
```

.pull-right[
```{r, echo=F, out.height = 400}
ggplot(iris) +
  aes(x = Sepal.Length) +
  geom_histogram(binwidth = 0.5,
                 fill = "coral3",
                 color = "goldenrod4")
```
]
---

class: inverse, middle, center

# Two Categorical Variables

---
# Stacked Bar-Plot
.pull-left[
```{r, eval=F}
ggplot(data = iris) +
  aes(x = Species, 
*      fill = Sepal.Length > 5) +
  geom_bar()
```
]
.pull-right[
```{r, echo=F, out.width=400}
ggplot(data = iris) +
  aes(x = Species, 
      fill = Sepal.Length > 5) +
  geom_bar() 
```
]
---
#Standardized Bar Plot
.pull-left[
```{r, eval=F}
ggplot(data = iris) +
  aes(x = Species, 
      fill = Sepal.Length > 5) +
*  geom_bar(position = "fill")
```
]
.pull-right[
```{r, echo=F, out.width=400}
ggplot(data = iris) +
  aes(x = Species, 
      fill = Sepal.Length > 5) +
  geom_bar(position = "fill")
```
]
---
class: middle
Those two plots looked the same... in which instance would they be different?

Hint: what is going on with the y-axis?

---
#Dodged Bar Plot
.pull-left[
```{r, eval=F}
ggplot(data = iris)  +
  aes(x = Species,
      fill = Sepal.Length > 5) + 
*  geom_bar(position = "dodge")
```
]
.pull-right[
```{r, echo=F, out.width=400}
ggplot(data = iris)  +
  aes(x = Species,
      fill = Sepal.Length > 5) + 
  geom_bar(position = "dodge")
```
]
---
class: inverse, middle, center
# Single Categorical and a Single Numerical Variable 

---
# Single Categorical, Single Numerical

.pull-left[
1. Create a ggplot using **iris** data frame
2. Map the **Species** to x-axis and **Sepal.Length** to y-axis
3. Add layer of violin plot
]
--
.pull-right[
```{r, eval = F}
ggplot(data = iris) +
  aes(x = Species,
      y = Sepal.Length)  +
*  geom_violin()
```
]

---

.pull-left[
```{r, echo = F, out.width=400, out.height=400}
ggplot(data = iris) +
  aes(x = Species,
      y = Sepal.Length)  +
  geom_violin() +
  labs(title = "Violin plot")
```
]

--

.pull-right[
```{r, echo=F, out.width=400, out.height=400}
ggplot(data = iris) +
  aes (x = Sepal.Length) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~Species)+
  labs(title = "Bar plot with facet wrap")
```
]

--

.pull-left[
```{r, eval=F}
ggplot(data = iris) +
  aes(x = Species,
      y = Sepal.Length)  +
  geom_violin() +
  labs(title = "Violin plot")
```
]

.pull-right[
```{r, eval=F}
ggplot(data = iris) +
  aes (x = Sepal.Length) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~Species)+
  labs(title = "Bar plot with facet wrap")
```
]
---
class:middle, center, inverse
#Two Numerical Variables
---
class: middle

.pull-left[
```{r, eval=F}
ggplot(iris) + 
  aes(x = Sepal.Length,
      y = Sepal.Width,
      color = Species) +
  geom_point()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris) + 
  aes(x = Sepal.Length,
      y = Sepal.Width,
      color = Species) +
  geom_point()
```
]
---
class:inverse, middle, center
#More than Two Categorical Variables
---
class: middle

.pull-left[
```{r, eval=F}
ggplot(iris) + 
  aes(x = Sepal.Length) +
  aes(y = Sepal.Width) +
  geom_point()
```
]
.pull-right[
```{r, echo=F}
ggplot(iris) + 
  aes(x = Sepal.Length) +
  aes(y = Sepal.Width) +
  geom_point()
```
]